---
title: 'Taking Questions: Week 12'
author: "Stats Corner"
date: "November 30, 2016"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
library(googlesheets)
library(dplyr)
library(reshape2)
library(tidyr)
library(knitr)
library(ggplot2)
library(shiny)
workbook <- gs_url("https://docs.google.com/spreadsheets/d/1c24qtCDF6MnL1I-nNG2ovymFB3fYj1NsWpLe3SGCbJs/pubhtml")

owner <- workbook %>% gs_read(ws = "Owner-Team Name")
games <- workbook %>% gs_read(ws = "Regular Season Games")

owner_games <- left_join(games,owner,by=c("year","team"))

data <- arrange(owner_games,game_id,points)
data$W <- rep(0,dim(data)[1])
data$L <- rep(0,dim(data)[1])
data$T <- rep(0,dim(data)[1])
data$PA <- rep(0,dim(data)[1])
data$result <- rep(" ",dim(data)[1])
data$last_of_streak <- rep(0,dim(data)[1])
data$opponent <- rep(" ",dim(data)[1])
games_played <- data %>% filter(points > 0)
### Loop to determine winners, and points against
for(i in seq(from=1,to=dim(games_played)[1],by=2)){
  
  games_played$PA[i]   <- games_played$points[i+1]
  games_played$PA[i+1] <- games_played$points[i]
  
  games_played$opponent[i] <- games_played$owner[i+1]
  games_played$opponent[i+1] <- games_played$owner[i]
  
  if(games_played$points[i] < games_played$points[i+1]){
    games_played$L[i] <- 1
    games_played$W[i+1] <- 1
    games_played$result[i] <- "L"
    games_played$result[i+1] <- "W"
    
  }
  if(games_played$points[i] == games_played$points[i+1]){
    games_played$T[i] <- 1
    games_played$T[i+1] <- 1
    games_played$result[i] <- "T"
    games_played$result[i+1] <- "T"
  }
  
}


games_played <- games_played %>%
  arrange(year,week,desc(points)) %>%
  group_by(year,week) %>%
  mutate(rk = rank(points,ties.method="min"),
         opp_rk = rank(PA,ties.method="min")) 

# Now to calculate and add up proportional wins
games_played$pw<- ifelse(games_played$year==2009,(games_played$rk-1)/9,(games_played$rk-1)/11)
games_played$opp_pw <- ifelse(games_played$year==2009,(games_played$opp_rk-1)/9,(games_played$opp_rk-1)/11)

season_summary <- games_played %>%
  filter(year != 2016) %>%
  ungroup() %>%
  mutate(luck = W - pw) %>%
  group_by(year,owner) %>%
  summarise(pw_sum = sum(pw),
            luck_sum = sum(luck),
            points_sum = sum(points),
            pa_sum = sum(PA),
            sos = mean(opp_pw)) %>%
  ungroup() %>%
  mutate(pw_rank = dense_rank(-pw_sum),
         pw_rank_btm = dense_rank(pw_sum),
         luck_rank_top = dense_rank(-luck_sum),
         luck_rank_btm = dense_rank(luck_sum),
         points_rank = dense_rank(-points_sum),
         pa_rank = dense_rank(-pa_sum),
         points_rank_btm = dense_rank(points_sum),
         sos_rank = dense_rank(-sos))

```


## "In honor of my 0.3 win over RJ, what is the smallest margin of victory we've had since the move to decimal scoring?"
Back in 2009 before we had decimal scoring we had two ties, Ready and Shokunbi tied at 97 in Week 4 then Kane and Ready tied in week 11 at 123.

In 2010 we went to decimal scoring and have had 12 games decided by less than a point, or almost two per season. 

```{r,echo=FALSE}
games_played %>%
  ungroup() %>%
  filter(year != 2009 & ((W == 1) | (T == 1 & owner != "Shokunbi"))) %>%
  mutate(Year = year,
         Week = week,
         Margin = points-PA,
         Winner = owner,
         Loser = opponent,
         Score = paste0(points," - ",PA)) %>%
  arrange(abs(Margin)) %>%
  ungroup() %>%
  slice(1:12) %>%
  select(Year,Week,Margin,Winner,Loser,Score) %>%
  kable(digits = 1, align = 'lrrrrr')
```

Also, we still had a tie! Ready and Shokunbi tied again at 131.0 in Week 4 of 2010 making my variable names sound misleading. Himmy and RJ's battle comes in at tied for 4th closest all time. Shout out to D'Skae for enduring a heartbreaking 0.1 point loss in 2011 to Matovina. 

## Record Chasing
```{r,echo=FALSE,include = FALSE}


twelve_weeks <- games_played %>% filter(week <= 12)
twelve_weeks_summary <- twelve_weeks %>%
  ungroup() %>%
  mutate(luck = W - pw) %>%
  group_by(year,owner) %>%
  summarise(pw_sum = sum(pw),
            luck_sum = sum(luck),
            pa_sum = sum(PA),
            points_sum = sum(points))

twelve_weeks_summary %>%
  ungroup() %>%
  arrange(-pw_sum) %>%
  slice(1:10) %>%
  left_join(.,season_summary %>% select(year,owner,pw_sum,pw_rank), by = c("year","owner")) %>%
  select(1,2,3,5,6) %>%
  `colnames<-`(c("Year", "Owner", "PW Through 12 Weeks", "PW Season Total","All Time Rank")) %>%
  kable(digits = 2)

```



### Hartman and Proportional Wins
```{r,echo=FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(-pw_sum) %>%
  slice(1:5) %>%
  left_join(.,season_summary %>% select(year,owner,pw_sum,pw_rank), by = c("year","owner")) %>%
  select(1,2,3,7,8) %>%
  `colnames<-`(c("Year", "Owner", "PW Through 12 Weeks", "PW Season Total","All Time Rank")) %>%
  kable(digits = 2)
```

Hartman came up short of 1 full PW last week so his shot at tying Ola's record is gone.

### Hartman and Total Points
```{r, echo = FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(-points_sum) %>%
  slice(1:5) %>%
  left_join(.,season_summary %>% select(year,owner,points_sum,points_rank), by = c("year","owner")) %>%
  select(1,2,6,7,8) %>%
  `colnames<-`(c("Year", "Owner", "Points Through 12 Weeks", "Points Season Total","All Time Rank")) %>%
  kable(digits = 2)
```

Hartman needs 153.4 points to claim the single season points record. 

### Olson and Luck
```{r, echo = FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(-luck_sum) %>%
  slice(1:7) %>%
  left_join(.,season_summary %>% select(year,owner,luck_sum,luck_rank_top), by = c("year","owner")) %>%
  select(1,2,4,7,8) %>%
  `colnames<-`(c("Year","Owner","Luck Through 12 Weeks","Season Total Luck","All Time Rank")) %>%
  kable(digits = 2)
```

With a good performance last week and a loss, Olson drops out of the Luck leaderboard. It is mathematically impossible for him to be the luckiest team of all time.

### Kane and (Un)Luck
```{r,echo = FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(luck_sum) %>%
  slice(1:5) %>%
  left_join(.,season_summary %>% select(year,owner,luck_sum,luck_rank_btm), by = c("year","owner")) %>%
  select(1,2,4,7,8) %>%
  `colnames<-`(c("Year","Owner","Luck Through 12 Weeks","Season Total Luck","All Time Rank")) %>%
  kable(digits = 2)
```

I put up a stinker last week so I couldn't become more unlucky. The single season record is in play, but my opponent is Ready and he has got a lot of guys (Cleveland Browns) on BYE this week so god help me if I don't win this one. 2012 Higdon can keep this record.


### Kane and Points Against
```{r,echo=FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(-pa_sum) %>%
  slice(1:5) %>%
  left_join(.,season_summary %>% select(year,owner,pa_sum,pa_rank), by = c("year","owner")) %>%
  select(1,2,5,7,8) %>%
  `colnames<-`(c("Year", "Owner", "Points Against Through 12 Weeks", "Points Against Season Total","All Time Rank")) %>%
  kable(digits = 2)
```

Again, I'm playing Ready and he is unlikely to score 140.8 and give me this terrible record. 2012 Higdon can keep this one too. 

### Harris, Ready and Fewest Total Points
```{r,echo = FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(points_sum) %>%
  slice(1:8) %>%
  left_join(.,season_summary %>% select(year,owner,points_sum,points_rank_btm), by = c("year","owner")) %>%
  select(1,2,6,7,8) %>%
  `colnames<-`(c("Year", "Owner", "Points Through 12 Weeks", "Points Season Total","All Time Rank")) %>%
  kable(digits = 2)
```

Harris will escape this record with a strong finish to the season. Ready likely will also. He needs to score 31.8 points to avoid this dubious distinction. But Josh McCown has a BYE so we don't know what's going to happen.   

### Harris, Ready and Fewest Proportional Wins
```{r, echo=FALSE}
twelve_weeks_summary %>%
  ungroup() %>%
  arrange(pw_sum) %>%
  slice(1:5) %>%
  left_join(.,season_summary %>% select(year,owner,pw_sum,pw_rank_btm), by = c("year","owner")) %>%
  select(1,2,3,7,8) %>%
  `colnames<-`(c("Year", "Owner", "PW Through 12 Weeks", "PW Season Total","All Time Rank")) %>%
  kable(digits = 2)

```

Last year's Thieneman team was an abomination. This year's Harris and Ready are close. I suspect Ken will manage his roster and escape this fate, he needs to score no lower than 10th overall in points to avoid at least a tie of the record. Ready needs to not score lower than 11th, but again his Cleveland Browns have a BYE and I'm not sure he can field a team. 