---
title: The Single Game Points Record Championship Belt
author: John Kane
date: '2017-11-28'
slug: 
categories:
  - 2017
  - Records
tags: 
  - 2017
  - Records
---


Welcome back to Stats Corner posts for the 2017 season! 

This is the first post on the new platform. The new platform allows for better graphics, including animations. It's in the spirit of, "I can so I will" that a bring you this post. 

```{r,echo=FALSE, include = FALSE}
library(googlesheets)
library(xml2)
library(httr)
library(curl)
library(shiny)
library(dplyr)
library(reshape2)
library(tidyr)
library(readr)
library(magrittr)
library(stringr)

workbook <- gs_url("https://docs.google.com/spreadsheets/d/1c24qtCDF6MnL1I-nNG2ovymFB3fYj1NsWpLe3SGCbJs/pubhtml")

owner <- workbook %>% gs_read(ws = "Owner-Team Name")
games <- workbook %>% gs_read(ws = "Regular Season Games")
playoff_games <- workbook %>% gs_read(ws = "Playoff Games")

owner_games <- left_join(games,owner,by=c("year","team"))
owner_games$week <- as.character(owner_games$week)
owner_games$game <- as.character(owner_games$game)
owner_games$game_id <- as.character(owner_games$game_id)
owner_playoff_games <- left_join(playoff_games,owner,by = c("year","team"))

all_games <- bind_rows(owner_games,owner_playoff_games)

## Want to know after every week, what is the top 10.
table(playoff_games$week)

all_games %<>%
  mutate(week_txt = ifelse(week == "R1","14",
                       ifelse(week == "R2","15",
                              ifelse(week == "R3","16",week_txt))))
         # ,
         # week = ifelse(week_txt == "14",14,
         #               ifelse(week_txt == "15",15,
         #                      ifelse(week_txt == "16",16,week))))


all_games$year_week <- paste0(all_games$year,all_games$week_txt)
#all_games$year_week <- paste0(all_games$year," Week ",all_games$week)
table(all_games$year_week)

all_games$team_year_week <- paste0(all_games$team,"-",all_games$year_week)

all_games$team_year_week

all_games$team_points <- paste0(all_games$team," - ",as.character(all_games$points))

f <- function(week_flag){
  output <-
  all_games %>%
    filter(year_week <= as.numeric(week_flag)) %>%
    arrange(desc(points)) %>%
    mutate(rank = row_number(),
           week = week_flag) %>%
     slice(1:10) %>%
    select(rank,year_week,week,team,points,owner,team_year_week,team_points)
  return(output)
}

f("201010")

weekly_top_10 <- lapply(list(unique(all_games$year_week)),function(x) lapply(x,f))
all_top_10 <- do.call(rbind,weekly_top_10[[1]])
head(all_top_10)

unique(all_games$year_week) 


year_week_mapping_df <- data.frame(year_week = unique(all_games$year_week)) %>%
  arrange(year_week) %>%
  mutate(obs = row_number())

all_top_10 %<>%
  inner_join(.,year_week_mapping_df, by = c("week" = "year_week"))

all_number1 <- filter(all_top_10,rank==1)
all_number1 %>%
  group_by(team_year_week,owner) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```


```{r}
summary <- all_number1 %>%
  group_by(owner,team_year_week,points) %>%
  summarise(weeks_at_no_1 = n(),
            first_week = min(as.numeric(week)),
            last_week = max(as.numeric(week))) %>%
  ungroup() %>%
  arrange(first_week) %>%
  mutate(first_week_char = paste0(str_sub(first_week,1,4)," Week ",str_replace(str_sub(first_week,5,6),'0','')),
         last_week_char = paste0(str_sub(last_week,1,4)," Week ",str_replace(str_sub(last_week,5,6),'0',''))) %>%
  select(1,2,3,4,7,8)


summary2 <- summary
summary2$start <- str_locate(summary$team_year_week,'-')[,1]
summary2$team <- str_sub(summary$team_year_week,1,summary2$start-1)
summary2 %>%
  select(1,8,3,4,5,6) %>%
  `colnames<-`(c("Owner", "Team", "Points","Weeks with Record","First Week","Last Week"))
```


I hope you enjoy like I did, seeing the names of teams from seasons past. 

Losman's Gamblers had a stranglehold on the record for two and half 


Try it with tweener
```{r}
library(tweenr)
tweenr_df <- all_top_10
tweenr_df$ease_func = 'linear'
tweenr_df$week <- as.numeric(tweenr_df$week)
head(tweenr_df)

all_top_10

att2 <- all_top_10

head(att2)
att2 %<>% select(rank,week,team,points,obs)

split_lists <- split(att2,att2$week)

dt <- tween_states(data = split_lists, tweenlength = 1, statelength = 1, ease = 'linear', nframes = 500)

head(all_top_10)

```


```{r}
library(ggplot2)
library(gganimate)
library(ggforce)
library(tweenr)

# Making up data
d <- data.frame(x = rnorm(20), y = rnorm(20), time = sample(100, 20), alpha = 0, 
                size = 1, ease = 'elastic-out', id = 1:20, 
                stringsAsFactors = FALSE)
d2 <- d
d2$time <- d$time + 10
d2$alpha <- 1
d2$size <- 3
d2$ease <- 'linear'
d3 <- d2
d3$time <- d2$time + sample(50:100, 20)
d3$size = 10
d3$ease <- 'bounce-out'
d4 <- d3
d4$y <- min(d$y) - 0.5
d4$size <- 2
d4$time <- d3$time + 10
d5 <- d4
d5$time <- max(d5$time)
df <- rbind(d, d2, d3, d4, d5)

# Using tweenr
dt <- tween_elements(df, 'time', 'id', 'ease', nframes = 500)

# Animate with gganimate
p <- ggplot(data = dt) + 
    geom_point(aes(x=x, y=y, size=size, alpha=alpha, frame = .frame)) + 
    scale_size(range = c(0.1, 20), guide = 'none') + 
    scale_alpha(range = c(0, 1), guide = 'none') #+ 
    ggforce::theme_no_axes()
animation::ani.options(interval = 1/24)
gganimate(p, 'dropping balls.gif', title_frame = F)
```


```{r}
plot3 <- ggplot(data = all_top_10,
               aes(x = obs, y = rank, frame = week)) +
  geom_path(aes(cumulative = TRUE, group = team)) +
  geom_label(aes(x=obs,y=rank,label = team_points)) +
  theme_minimal() +
  scale_y_reverse() +
  scale_x_continuous("Time",
                     lim = c(-35,175),
                     breaks = seq(from = 1, to = 16*9,by=16),
                     labels = seq(from = 2009, to = 2017)) +
  ggtitle("The Single Game Points Record Championship Belt \n")

library(tweenr)

all_top_10_2 <- all_top_10 %>%
  arrange(team,obs) %>%
  #select(obs,rank,team,week,team_points) %>%
  select(obs,rank,team_points) %>%
  rename(x = obs, y = rank, time = obs, group = team_points) %>%
  mutate(ease = "linear")

all_top_10_2$group

is.vector(all_top_10_2$group)
is.vector(all_top_10$team_points)

all_top_10_tween <- tween_elements(data = all_top_10_2,'time','group','ease',nframes = 276) #138*5

library(ggplot2)
library(gganimate)
plot_it <- ggplot(data = all_top_10_tween, aes(x=time,y=y,frame = .frame, group = .group, cumulative = T,label = .group)) +
  geom_path() +
  geom_label() +
  theme_minimal() +
  scale_y_reverse() +
  scale_x_continuous("Time",
                     lim = c(-35,175),
                     breaks = seq(from = 1, to = 16*9,by=16),
                     labels = seq(from = 2009, to = 2017)) +
  ggtitle("The Single Game Points Record Championship Belt \n")

gganimate(plot_it, interval = 0.2)


```




