---
title: "Stats Corner"
output: html_document
---
contact: stats.corner@gmail.com

Welcome back to Stats Corner for the 2015 Season!

This place has a new look, a new location on the internet and new ideas. I will no longer be restricted to message board posts. Bear with me as the style evolves. Here in Week 1 I want to focus on content and content only. Future posts should have a better style.

## Scoring the Draft

Coming out of Draft Day I'm sure most of you were feeling pretty good. Minus D'Skae of course, he never likes his team. Can we quantify how well everyone's draft was? Of course we can. It's pretty straightforward: we cull the projections from various sites and join them up to everyone's roster. Recall that ESPN's projections aren't great. For perspective, I also looked how CBSsportsline.com and Yahoo! projected players to perform.

#### ~~Quick~~ notes on methodology
- I assume you're playing a QB at OP. God help you if you're not.
- I assume head coaches will score you 0 all season (probably not true).
- I'm only scoring your starters: 2 QB's, 2 RB's, 3 WR's, 1 TE, 1 D/ST and 1 K. 
- ESPN let me download projections unique to our league's scoring. CBS and Yahoo did not. Therefore interpret their D/ST and K projections loosely. Still informative, but probably wrong. 
- I assume no halfback passes. I do assume my #1 WR does [this](https://www.youtube.com/watch?v=bvS9fYHBRvo) again though. Just kidding. 
- No return touchdowns count toward individual players.
- For CBS projections, no rushing yards by a WR. 
- RJ, I'm sorry but your drafting of David Cobb (he plays for the Titans everyone, including you RJ) was not counted. He didn't appear in all draft projections so I had to leave him out. ESPN had him at 26.6 points for the __year__ so hopefully he wasn't going to be your RB1. Or RB4.
- Matovina you drafted Caleb Sturgis. Caleb Sturgis is not currently on an NFL roster. So I had to impute your Kicker projections with the level one unit below the next-to-worst kicker on the other 11 teams. Though you won't see it I named him "Replacement Level."

### ESPN

Without further adieu, the results from ESPN. I threw in Z-scores because they'll be useful later.


```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(sqldf)
setwd('/home/john/stats_corner/2015/draft_scoring')
final_projections <- read.csv("final_projections_data.csv",stringsAsFactors = FALSE)

qbs_espn <- filter(final_projections,
                   pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(2,espn_points)
rbs_espn <- filter(final_projections,
                   pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(2,espn_points)
wrs_espn <- filter(final_projections,
                   pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(3,espn_points)
tes_espn <- filter(final_projections,
                   pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(1,espn_points)
ks_espn <- filter(final_projections,
                   pos == 'K') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(1,espn_points)
matovina_k <- data.frame(owner = 'Matovina',
                         player = 'Replacement Level',
                         pos = 'K',
                         espn_points = 129)
ks_espn <- bind_rows(ks_espn,matovina_k)
dst_espn <- filter(final_projections,
                   pos == 'D/ST') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    top_n(1,espn_points)
espn_starters <- bind_rows(qbs_espn,
                           rbs_espn,
                           wrs_espn,
                           tes_espn,
                           ks_espn,
                           dst_espn)
summarise_espn <- espn_starters %>%
    group_by(owner) %>%
    summarise(total_points = sum(espn_points)) %>%
    arrange(desc(total_points))
summarise_espn$z <- (summarise_espn$total_points - mean(summarise_espn$total_points))/sd(summarise_espn$total_points)
print(summarise_espn)
```

Before you go erasing names in the record book for most points in a season, note how just about every team is projected to set a new record. These are *probably* overestimates. Still fun though. Poor Ola. 

### What about Yahoo! and CBSsportsline.com?
Because ESPN doesn't have a great track record at making projections, let's look at two other sources. It's not that Yahoo! and CBSsportsline.com are any better necessarily, but a composite view has been shown to be the most accurate. Because of things previously mentioned I'll use the z-scores for each of the projection systems to make things an even playing field. The final metric will be the sum of three z-scores. It's essentially meaningless but useful for the rankings. 

```{r, echo=FALSE, message = FALSE}
qbs_yahoo <- filter(final_projections,
                  pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(2,yahoo_points)
rbs_yahoo <- filter(final_projections,
                  pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(2,yahoo_points)
wrs_yahoo <- filter(final_projections,
                  pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(3,yahoo_points)
tes_yahoo <- filter(final_projections,
                  pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(1,yahoo_points)
ks_yahoo <- filter(final_projections,
                 pos == 'K') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(1,yahoo_points)
matovina_k <- data.frame(owner = 'Matovina',
                         player = 'Replacement Level',
                         pos = 'K',
                         yahoo_points = 112.5)
ks_yahoo <- bind_rows(ks_yahoo,matovina_k)
dst_yahoo <- filter(final_projections,
                  pos == 'D/ST') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    top_n(1,yahoo_points)
yahoo_starters <- bind_rows(qbs_yahoo,
                          rbs_yahoo,
                          wrs_yahoo,
                          tes_yahoo,
                          ks_yahoo,
                          dst_yahoo)
summarise_yahoo <- yahoo_starters %>%
    group_by(owner) %>%
    summarise(total_points = sum(yahoo_points)) %>%
    arrange(desc(total_points))
summarise_yahoo$z <- (summarise_yahoo$total_points - mean(summarise_yahoo$total_points))/sd(summarise_yahoo$total_points)
qbs_cbs <- filter(final_projections,
                   pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(2,cbs_points)
rbs_cbs <- filter(final_projections,
                   pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(2,cbs_points)
wrs_cbs <- filter(final_projections,
                   pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(3,cbs_points)
tes_cbs <- filter(final_projections,
                   pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(1,cbs_points)
ks_cbs <- filter(final_projections,
                  pos == 'K') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(1,cbs_points)
matovina_k <- data.frame(owner = 'Matovina',
                         player = 'Replacement Level',
                         pos = 'K',
                         cbs_points = 129.2)
ks_cbs <- bind_rows(ks_cbs,matovina_k)
dst_cbs <- filter(final_projections,
                   pos == 'D/ST') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    top_n(1,cbs_points)
cbs_starters <- bind_rows(qbs_cbs,
                           rbs_cbs,
                           wrs_cbs,
                           tes_cbs,
                           ks_cbs,
                           dst_cbs)
summarise_cbs <- cbs_starters %>%
    group_by(owner) %>%
    summarise(total_points = sum(cbs_points)) %>%
    arrange(desc(total_points))
summarise_cbs$z <- (summarise_cbs$total_points - mean(summarise_cbs$total_points))/sd(summarise_cbs$total_points)


all_scores <- sqldf('
                    select
                        a.owner
                    ,   round(a.total_points) as espn_points
                    ,   round(a.z,3) as espn_z
                    ,   round(b.total_points) as cbs_points
                    ,   round(b.z,3) as cbs_z
                    ,   round(c.total_points) as yahoo_points
                    ,   round(c.z,3) as yahoo_z
                    from
                        summarise_espn as a
                    ,   summarise_cbs as b
                    ,   summarise_yahoo as c
                    where
                        a.owner = b.owner
                    and a.owner = c.owner')
all_scores <- all_scores %>% mutate(total_z = round(espn_z + cbs_z + yahoo_z,3)) %>% arrange(desc(total_z)) %>% select(owner, espn_z, yahoo_z, cbs_z, total_z)
print(all_scores)
```

Well look what we have here! None other than Matovina coming out on top. In fact, his starters rated out first overall in *both* Yahoo! and CBSsportsline.com rankings. __And this is with a replacement level kicker everyone__. Look out for Matovina, may he stay healthy.

### Scoring the Bench Warmers
> "I like depth." - Brenden Regan

Fair enough Regan, and I suppose I can't argue with that simple, declarative sentence from draft night. But you can only play 10 and good for you if your 4th WR is a stud, you only get points for your top 3. Hopefully you play the right guy.

It wouldn't be fair to simply add up bench points. People load up their rosters in different ways. For example, Ola attempted to draft anyone with an arm in the draft, whereas others went without a backup QB (gulp). I'll take your top backup at each position. I do not consider kickers (no one has a backup) nor D/ST (only one team drafted a backup. Good on you, McShane). 


If you happened to not draft a backup QB, RB, WR or TE then I award you no points, and may God have mercy on your soul. 

Here they are, the bench rankings!
```{r, echo=FALSE, message=FALSE}
setwd('/home/john/stats_corner/2015/draft_scoring')
final_projections <- read.csv("final_projections_data.csv",stringsAsFactors = FALSE)

qbs_espn_bench <- filter(final_projections,
                   pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
rbs_espn_bench <- filter(final_projections,
                   pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
wrs_espn_bench <- filter(final_projections,
                   pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 4)
tes_espn_bench <- filter(final_projections,
                   pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(espn_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 2)

espn_bench <- bind_rows(qbs_espn_bench,
                           rbs_espn_bench,
                           wrs_espn_bench,
                           tes_espn_bench)
summarise_espn_bench <- espn_bench %>%
    group_by(owner) %>%
    summarise(total_points = sum(espn_points)) %>%
    arrange(desc(total_points))
summarise_espn_bench$z <- (summarise_espn_bench$total_points - mean(summarise_espn_bench$total_points))/sd(summarise_espn_bench$total_points)
qbs_cbs_bench <- filter(final_projections,
                         pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
rbs_cbs_bench <- filter(final_projections,
                         pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
wrs_cbs_bench <- filter(final_projections,
                         pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 4)
tes_cbs_bench <- filter(final_projections,
                         pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(cbs_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 2)

cbs_bench <- bind_rows(qbs_cbs_bench,
                        rbs_cbs_bench,
                        wrs_cbs_bench,
                        tes_cbs_bench)
summarise_cbs_bench <- cbs_bench %>%
    group_by(owner) %>%
    summarise(total_points = sum(cbs_points)) %>%
    arrange(desc(total_points))
summarise_cbs_bench$z <- (summarise_cbs_bench$total_points - mean(summarise_cbs_bench$total_points))/sd(summarise_cbs_bench$total_points)
qbs_yahoo_bench <- filter(final_projections,
                         pos == 'QB') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
rbs_yahoo_bench <- filter(final_projections,
                         pos == 'RB') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 3)
wrs_yahoo_bench <- filter(final_projections,
                         pos == 'WR') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 4)
tes_yahoo_bench <- filter(final_projections,
                         pos == 'TE') %>%
    group_by(owner) %>%
    arrange(desc(yahoo_points)) %>%
    mutate(n = row_number()) %>%
    filter(n == 2)

yahoo_bench <- bind_rows(qbs_yahoo_bench,
                        rbs_yahoo_bench,
                        wrs_yahoo_bench,
                        tes_yahoo_bench)
summarise_yahoo_bench <- yahoo_bench %>%
    group_by(owner) %>%
    summarise(total_points = sum(yahoo_points)) %>%
    arrange(desc(total_points))
summarise_yahoo_bench$z <- (summarise_yahoo_bench$total_points - mean(summarise_yahoo_bench$total_points))/sd(summarise_yahoo_bench$total_points)


all_scores_bench <- sqldf('
                    select
                    a.owner
                    ,   round(a.z,3) as espn_z
                    ,   round(b.z,3) as cbs_z
                    ,   round(c.z,3) as yahoo_z
                    from
                    summarise_espn_bench as a
                    ,   summarise_cbs_bench as b
                    ,   summarise_yahoo_bench as c
                    where
                    a.owner = b.owner
                    and a.owner = c.owner')
all_scores_bench <- all_scores_bench %>%
    mutate(total_z = round(espn_z + cbs_z + yahoo_z,3)) %>%
    arrange(desc(total_z)) %>%
    print()
```

Well now Matovina *really* better stay healhty. His bench is awful. Thieneman and Ready enjoy that combo of being bottom 3 in starters and bottom 3 in bench. I say look out for Himmy: third ranked starters and 4th ranked bench. 

Good luck to everyone this season, remember this analysis is not your destiny. And with Ola's weak starting lineup and strong bench be expecting maybe a few more trade texts than usual. 