---
title: 2019 Regular Season Recap
author: John Kane
date: '2019-12-11'
slug: 2019-regular-season-recap
categories: []
tags: []
cover: /images/cover.jpg
---


```{r}
library(googlesheets)
library(xml2)
library(httr)
library(curl)
library(shiny)
library(dplyr)
library(reshape2)
library(tidyr)
library(readr)
library(magrittr)
library(stringr)
library(ggplot2)
library(gganimate)
library(tweenr)
library(knitr)
library(ggrepel)

workbook <- gs_url("https://docs.google.com/spreadsheets/d/1c24qtCDF6MnL1I-nNG2ovymFB3fYj1NsWpLe3SGCbJs/pubhtml")

owner <- workbook %>% gs_read(ws = "Owner-Team Name")
games <- workbook %>% gs_read(ws = "Regular Season Games")

owner_games <- left_join(games,owner,by=c("year","team"))
owner_games$week <- as.character(owner_games$week)
owner_games$game <- as.character(owner_games$game)
owner_games$game_id <- as.character(owner_games$game_id)


data <- arrange(owner_games,game_id,points)
data$W <- rep(0,dim(data)[1])
data$L <- rep(0,dim(data)[1])
data$T <- rep(0,dim(data)[1])
data$PA <- rep(0,dim(data)[1])
data$result <- rep(" ",dim(data)[1])
data$opponent <- rep(" ",dim(data)[1])
games_played <- data %>% filter(points > 0)
### Loop to determine winners, and points against
for(i in seq(from=1,to=dim(games_played)[1],by=2)){
  
  games_played$PA[i]   <- games_played$points[i+1]
  games_played$PA[i+1] <- games_played$points[i]
  
  games_played$opponent[i] <- games_played$owner[i+1]
  games_played$opponent[i+1] <- games_played$owner[i]
  
  if(games_played$points[i] < games_played$points[i+1]){
    games_played$L[i] <- 1
    games_played$W[i+1] <- 1
    games_played$result[i] <- "L"
    games_played$result[i+1] <- "W"
    
  }
  if(games_played$points[i] == games_played$points[i+1]){
    games_played$T[i] <- 1
    games_played$T[i+1] <- 1
    games_played$result[i] <- "T"
    games_played$result[i+1] <- "T"
  }
  
}


games_played <- games_played %>%
  arrange(year,week,desc(points)) %>%
  group_by(year,week) %>%
  mutate(rk = rank(points,ties.method="min"),
         opp_rk = rank(PA,ties.method="min")) 

# Now to calculate and add up proportional wins
games_played$pw<- ifelse(games_played$year==2009,(games_played$rk-1)/9,(games_played$rk-1)/11)
games_played$opp_pw <- ifelse(games_played$year==2009,(games_played$opp_rk-1)/9,(games_played$opp_rk-1)/11)
games_played$year_owner <- paste0(games_played$year," ",games_played$owner)
```

```{r}
data19 <- filter(games_played, year == 2019)

data19_standings <-
  data19 %>%
  group_by(team) %>%
  summarise(wins = sum(W),
            pw = sum(pw)) %>%
  arrange(desc(points))

data19_standings
```


The 2019 Regular Season has come and gone.


I'll do a brief stat for each of the 12 teams. Starting with the 12 seed down to the first.



### 12. No One Drafts Like Papa (1-11, 2.0 PW)

There is no where to go but up for Joe D. in 2020. 

* 1 victory (fewest ever)
* 2.0 PW (fewest ever)
* 1058.6 points (second fewest ever, (2010 Thieneman))

This team is probably the worst in league history. Haven't seen a team this bad since the early part of the decade with Matovina and Thieneman.

### 11. Tough ACTin TinACTin (4-9, 5.7 PW) 


### 10. RJ's Team RJT3 (5-8, 5.9 PW)

### 9. Boats and RPOs (5-8, 6.6 PW)

### 8. Urban Fervor (6-7. 8.3 PW)

### 7. Whats a ZJ? (6-7, 6.8 PW)

### 6. Loafs n' Jugs (6-7, 5.7 PW)

### 5. 3 Blind Mice (8-5, 6.6 PW)

### 4. Joe Buck Yourself (8-5, 7.5 PW)

### 3. Dinosaur Train (9-4, 7.0 PW)

### 2. Viking Land Bound (10-3, 7.6 PW)

### 1. Anchor Arms (10-3, 8.1 PW)



